Drought and its Effects on Economy

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation2.tex
adding files, initial

\title{Drought and its Impacts on Economy in East Africa}  
\author{Monika Novackova} 
\date{December 2017} 
\institute{University of Sussex}

@@

\begin{frame}
\titlepage
\end{frame} `

@@

\section{Defining drought}

@@

\section{Drought indices}

@@

 \begin{itemize}
 \item \textbf{Standardized Precipitation Evapotranspiration Index} (SPEI, \citealp{SPEI})
\begin{itemize}
\item Extention of SPI
\item Accounts for potential evapotranspiration (hence captures impacts of increased temperature on water demand)
\end{itemize}
\item Number of other drought indicators and indices exist
\begin{itemize}
\item E.g.: Percent of Normal Precipitation, Drought Area Index (DAI), Soil Moisture Anomaly (SMA), Standardized Water-level Index (SWI), Normalized Difference Vegetation Index (NDVI)
\end{itemize}
\item  \citealp{monacelli2005, svoboda2016, zargar2011}: Overview of number of drought indices

\item Quantification and evaluation of drought indices for each of the three physical types of drought \citep{keyantash2002}
\end{itemize}

@@

\section{Effects of Drought on Economy}

@@

\frametitle{Effects of drought on economy}\label{Effects} 
\begin{block}{\textbf{Generalized Computable Equilibrium (CGE)} Models }

\begin{itemize}
\item \underline{\textbf{\cite{OxfamIDS}}}
\begin{itemize}
\item Exploring range of scenarios for food price increase in 2030
\begin{itemize}
\item 1. Baseline 2. Climate change 3. Climate change with adaptation 4. Adaptation only in sub-Saharan Africa
\end{itemize}
\item Global coverage, set of individual country models, linked through international trade
\item Effects of climate change (incl. drought) modelled as changes in factor productivity (usually negative)

\end{itemize}
\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item Ethiopia
\item Social Accounting Matrix (SAM)
\end{itemize}
\end{itemize}
\end{block}
\end{frame}

\note{Computable general equilibrium (CGE) models are a class of economic models that use actual economic data to estimate how an economy might react to changes in policy, technology or other external factors.\\
\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
\vspace{0.3cm}a\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a

\begin{frame}{\small Recent Relavant Literature}\label{Previous research}
\Wider[2em]{
\vspace{0.2cm}
\linespread{0.9}
\begin{itemize}
\small

\item \textbf{\cite{robinson2010}}: CGE
\begin{itemize}
\item Ethiopia. Udelat tabulku, v jednom sloupci co je fixed, v jednom co determined by model
\item ktere promenne se tam dosadi a ktere vylezou
\end{itemize}
\item \textbf{\cite{OxfamIDS}} CGE global
\begin{itemize}
\item probably also describe this CGE, how does it work, inputs, outputs
maybe it is like not that production is exogenous and price endo. but interdependent. and we have production shock as input $\rightarrow$ it gives price shock as output $\rightarrow$ it has again effect on production volumes
\end{itemize}

\item \textbf{\cite{Carlsson2013}}
\begin{itemize}
\item Analyse preferences of distributing the burden of reducing CO\textsubscript{2}
\item Higher educated people have higher WTP for distribution that is less costly for their country
\end{itemize}

\end{itemize}}
\end{frame}

\begin{frame}\label{Data3}
\frametitle{\small Data}

\vspace{-0.3cm}
      \begin{columns}
        \column{.15\linewidth}
        n=$4592$
              \column{.85\linewidth}
\begin{table}
\footnotesize
\label{Demographics} 
\begin{tabular}{llllllll} 
\hline
\rowcolor{PaleGreen} \textbf{Age} & 20 & 30 & 40 & 50 & 60 & 70 & 80 \\
 \hline
 \textbf{Frequency}  &$805$ & $1113$ & $812$ & $791$ & $657$ & $361$ & $53$\\
 \hline
\end{tabular}
\end{table}
 \end{columns}

      \begin{columns}
        \column{.335\linewidth}
\begin{table}
\footnotesize
\label{Sex} 
\vspace{-0.7cm}
\begin{tabular}{lll} 
\hline
\rowcolor{PaleGreen} \textbf{Gender} & \textbf{Freq.}  \\
 \hline
 Male & $2250$ \\
 Female& $2310$ \\
 \hline
\end{tabular}
\end{table}

         \column{.65\linewidth}
         \vspace{-1.1cm}
         \begin{figure}
         \caption*{\footnotesize{\vspace{-0.05cm} Education}}
           \vspace{-0.2cm}
 % \includegraphics[width = 0.9\textwidth]{Education.png} 
    \end{figure}
      \end{columns}

 \begin{figure}
 \vspace{-1.2cm}

 \caption*{\footnotesize{\vspace{-0.2cm}  \hspace{-2cm} Occupation}}
  \hspace{-2.3cm}
  \vspace{-0.55cm}  
%\includegraphics[width = 0.6\textwidth]{Occupation.png}  
\end{figure}

@@

\section{Methods}

\linespread{1}

\note{2 hlavni skupiny prediktoru: demographics a tak, a popisne...
 \\
\vspace{1cm}Nemela jsem moc hypotezu a literatura rika ruzne veci, v kazdem clanku neco jineho. takze selection methods...}

@@

\frametitle{Methods}\label{MethodsLasso} 
\textbf{Lasso} = Least Absolute Shrinkage and Selection Operator
\begin{itemize}
\item In literature preferred to stepwise procedures (Tibshirani, 1996; Efron et. al, 2003; Yuan and Lin, 2004; Shah, 2012) 
\begin{equation}\label{Lasso}
\begin{array}{lcll}

\widehat{\beta}(\lambda)&=& \underset{\beta}{arg~min} \bigg( \sum_{i=1}^{n}(Y_i - \boldsymbol{X} \beta)_i \bigg)^2 + \lambda \sum_{j=1}^p|\beta_j|, \\
\end{array}
\end{equation}

\begin{itemize}
\item $i$ ... index of observations
\item $j$ ... index of explanatory variables
\item $\lambda \geq 0$...penalty parameter
\end{itemize}

\item The estimator does variable selection in the sense that $\widehat{\beta_j}(\lambda) = 0 $ for some $j$'s
\item Cross-validation of a series of models with different values of $\lambda$
\item I choose the largest value of $\lambda$ such that the mean square error is within 1 standard error of the minimum MSE
\end{itemize}
\end{frame}

\note{\tiny
Why lasso prefered to stepwise??

why stepwise bad:
Efron at all: Stepwise can eliminate a good valid predictor in the second step if it happens to be correlted with the first selected predictor \\

http://www.stata.com/support/faqs/statistics/stepwise-regression-problems/\\

 It yields R-squared values that are badly biased to be high.\\
The F and chi-squared tests quoted next to each variable on the printout do not have the claimed distribution.\\
The method yields confidence intervals for effects and predicted values that are falsely narrow; see Altman and Andersen (1989).\\
It yields p-values that do not have the proper meaning, and the proper correction for them is a difficult problem.\\
It gives biased regression coefficients that need shrinkage (the coefficients for remaining variables are too large; see Tibshirani [1996]).\\
It has severe problems in the presence of collinearity.\\
It is based on methods (e.g., F tests for nested models) that were intended to be used to test prespecified hypotheses. \\

“The degree of correlation between the predictor variables affected the frequency with which authentic predictor variables found their way into the final model.” \\
 p is the number of parameters	 \\
\hspace{1cm} the selection depends on lambda
\hspace{1cm} R cross validates various models for series of different lamdas.
\hspace{1cm} }

@@

\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc ok}

\section{Results-Knowledge}\label{Results-Knowledge}

\section{Results-WTP}

\begin{frame}[plain]
\Wider[3.5em]{\vspace{-0.5cm}
\frametitle{\small Results: Willingness to pay}\label{ResultsWTP1} 
\begin{table}
%  bodie38  bodie70 Dunquan98  Dunquan80
\label{WTP1} 
\begin{footnotesize}

\begin{tabular}{m{0.2cm}m{0.05cm}|l|cc|ccc|} 
\hline
\rowcolor{PaleGreen} \multicolumn{3}{c|}{\textbf{OLS after selection by lasso}} &  \multicolumn{2}{c|}{\textbf{Tax gas}} &\multicolumn{2}{c}{\textbf{Tax fuel}} \\
\rowcolor{PaleGreen} \multicolumn{3}{c|}{Robust std. errors HC\textsubscript{0}, n=4178}& \textbf{(\RN{1})}&\textbf{(\RN{2})} & \textbf{(\RN{1})}&\textbf{(\RN{2}) } \\
\hline
 & &Female& + & +& -&+ \\ 
& & Age& -  *** & -  ***& -  ***& -  ***  \\ 
& & Debts& +  *** &+  ***&  +  ***&  +  *** \\ 
\RowColor \cellcolor{white} \multirow{-5}{*}{\hspace{-0.5cm} \vspace{-0.5cm} \rotatebox[origin=c]{55}{\textbf{\scriptsize{Demographics}}}}  &\cellcolor{white} & \cellcolor{white}Occupation & \multicolumn{2}{c}{\cellcolor{white} vary} &  \multicolumn{2}{|c}{ \textcolor{red}{Not included}    } \\
\hline
& & Understands interest & -  *** &-  ***  &-  ** &-  **\\ 
& & Understands inflation & -  *** &-  ***  &-  ***&-  ***\\

 \RowColor \cellcolor{white} \multirow{-3}{*}{\hspace{-0.3cm}\rotatebox[origin=c]{60}{\textbf{\scriptsize{Numeracy}}}} & \cellcolor{white} & \cellcolor{white}Cognitive reflection test&\multicolumn{2}{c}{\textcolor{red}{Not included}} &  \cellcolor{white} -   **&  \cellcolor{white} -   **\\

\hline

&  &Not getting fair share& \multirow{2}{*}{- ***}& \multirow{2}{*}{- ***} &\multirow{2}{*}{- **} & \multirow{2}{*}{- **}\\ 
& & \hspace{0.3cm} of nation's wealth& & \\ 
& & Hard work is important& -  ***& -  ***& -  ***& -  ***\\ 
\vspace{-0.2cm}& & Parents + kids have same&\multirow{2}{*}{+ ***}&  \multirow{2}{*}{+ ***} &  \multirow{2}{*}{+ ***}&  \multirow{2}{*}{+ ***} \\ 
\multirow{-5}{*}{\rotatebox[origin=c]{90}{\textbf{\textbf{Attitudes,}}}} & \hspace{-0.5cm} \multirow{-5}{*}{\rotatebox[origin=c]{90}{\textbf{\textbf{beliefs}}}} & \hspace{0.1cm} living standards as me& & \\

\hline

\RowColor \cellcolor{white} &\cellcolor{white} &\cellcolor{white}Clim. change risk perception&  \cellcolor{white}+ *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}&  \cellcolor{white}+ *** &\textcolor{red}{Not incl.}\\

\RowColor \cellcolor{white} &\cellcolor{white}&\cellcolor{white}Clim. change vs policies risk&  \cellcolor{white}+ *** & \cellcolor{white}+ *** &  \cellcolor{white}+ *** &  \cellcolor{white}+ ***\\

\RowColor \cellcolor{white}    & \cellcolor{white}    &  \cellcolor{white}How much is gas/fuel tax & \cellcolor{white} + *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}& \cellcolor{white} + *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}\\

 \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{Climate}}}     &  \hspace{-0.5cm} \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{change}}}    &  Female $\times$ Clim. policies & - ** & - * &  - & -\\

\hline
\RowColor \cellcolor{white} & \cellcolor{white} & \cellcolor{white}Inequity aversion index& \cellcolor{white} \scriptsize + {\tiny then} -  ***& \multicolumn{1}{c}{\textcolor{red}{Not incl.}} &\cellcolor{white} \scriptsize + {\tiny then} -  ***& \multicolumn{1}{c}{\textcolor{red}{Not incl.}} \\ 
\RowColor \cellcolor{white}&\cellcolor{white}&\cellcolor{white}Discount rate yr. 0 vs.1 &\cellcolor{white} + * &\multicolumn{1}{c}{\textcolor{red}{Not incl.}} &\cellcolor{white} + $\bullet$ &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}\\
&  & Consistent discount rate & + *** &+ ***&+ ***&+ *** \\
\RowColor \cellcolor{white} \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{\scriptsize{Behavioral}}}}& \cellcolor{white} &\cellcolor{white}Egalitarian & \multicolumn{2}{c}{\textcolor{red}{Not included}} & \cellcolor{white}- ***& \cellcolor{white}-*** \\
\hline
\end{tabular} 
\end{footnotesize}
\end{table} }
%checked 31.3.2017
\end{frame}

\note{Consistent answers within investments==0 consdr \\
climpol a climcare popsat otazky, mozna udelat o nich slide
\\

occupation: homemaker unemployed negative, otherwise insignificant \\

 to je taky duvod, proc si myslim, ze numeracy a literacy...ze to muze byt v politickych nazorech?? ze levicaci nemaji technicke vzdelani???}

\begin{frame}
% bodie53, Dunquan68lm
\frametitle{Results: Willingness to pay - robustness}\label{ResultsLassoGas}

\begin{itemize}
\item OLS after selection by lasso
\item With additional control variables
\end{itemize}

\begin{table}
\label{LassoBigSample} 
\begin{small}
\begin{tabular}{l|cc|cc} 
\hline
\rowcolor{PaleGreen} Dependent variable: &  \multicolumn{2}{c|}{\textbf{Tax gas}} &\multicolumn{2}{c}{\textbf{Tax fuel}} \\
\hline
Demographics & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust} \\ 
Numeracy & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
Attitudes, beliefs & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
Climate change & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
Behavioral & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
\hline

  Income & -& & - &**\\
    Climate knowledge &- & $\bullet$ & +\\
Social value orientation & + & &+&*\\

 Risk aversion Kehneman-Tversky & - & & + & \\
  Government should redistribute income & + &$\bullet$ &+& $\bullet$\\
\hline
 \multicolumn{1}{l|}{Robust standard errors HC\textsubscript{0}}&  \multicolumn{2}{l|}{n=$4178$}&  \multicolumn{2}{l}{n=$4184$}\\
\end{tabular} 
\end{small}
\end{table}

\note{ redistribute jsem dala jako politickou. Mozna ukazat bar??

         \textbf{Average preferred tax rate for \colorbox{babypink}{gas}} grouped by:
\begin{itemize}       
    \item    \textbf{\small Understands compound interest}

     \begin{figure}
      \begin{columns}
        \column{.4\linewidth}

  \caption*{ \hspace{0.6cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.6cm} \textcolor{red}{$\blacksquare$} Female
  \\ \vspace{0.2cm} \hspace{0.6cm} $1$ = Understands}
        \column{.6\linewidth}

 %   \vspace{-0.2cm}   \hspace{0.6cm}   \includegraphics[width = 0.8\textwidth]{GasInterest.png}  %565*238
 \vspace{-0.3cm}   \caption*{\footnotesize \hspace{0.6cm}$0$ \hspace{1cm}$0.5$ \hspace{1cm}$1$ }
      \end{columns}
    \end{figure}

\vspace{-0.3cm}
    \item    \textbf{\small Hard work is important}
\end{itemize}

\vspace{-0.2cm}
    \begin{figure}
      \begin{columns}
        \column{.5\linewidth }
        \vspace{-1.5cm}
  \caption*{ \hspace{0.8cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.8cm} \textcolor{red}{$\blacksquare$} Female
  \\ \vspace{0.2cm} \hspace{0.8cm} $1$ = Not important at all  \\ \hspace{0.8cm} $5$ = Essential}
  \vspace{-1.5cm}        
        \column{.5\linewidth}

  % \includegraphics[width = 0.85\textwidth]{GasHardwork.png}  %565*238
 \vspace{-0.3cm}   \caption*{\footnotesize  $1$ \hspace{0.5cm}$2$ \hspace{0.5cm}$3$ \hspace{0.5cm}$4$ \hspace{0.5cm}$5$  }
      \end{columns}
    \end{figure}

@@

    \textbf{Average preferred tax rate for \colorbox{babypink}{fuel}} grouped by:
    \begin{itemize}       
    \item \textbf{\small Cognitive Reflection Test}

    \begin{figure}
      \begin{columns}
        \column{.5\linewidth}
  \caption*{ \hspace{0.4cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.7cm} \textcolor{red}{$\blacksquare$} Female}
        \column{.5\linewidth}

  %    \vspace{-0.5cm}  \includegraphics[width = 1\textwidth]{FuelNumeracy.png}  %565*238
 \vspace{-0.3cm}   \caption*{\footnotesize \hspace{-4.5cm} Number of correct answers:\hspace{1.1cm} $0$\hspace{0.3cm}$0.5$ \hspace{0.2cm}$1$ \hspace{0.2cm}$1.5$ \hspace{0.2cm}$2$ \hspace{0.2cm}$2.5$\hspace{0.2cm}$3$ }
      \end{columns}
    \end{figure}

\vspace{-0.15cm} 
 \item \textbf{\small Ordinary working people do no get their fair share of the nation's wealth}
\end{itemize}
\vspace{-0.5cm}
  \begin{figure}
      \begin{columns}

        \column{.5\linewidth}
  \caption*{ \hspace{0.8cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.8cm} \textcolor{red}{$\blacksquare$} Female
  \\ \vspace{0.2cm} \hspace{0.8cm} $-2$ = Strongly disagree  \\ \hspace{1.1cm} $2$ = Strongly agree}
\vspace{-1cm}        
        \column{.5\linewidth}
   \vspace{-0.2cm}     \hspace{-0.8cm} 
  % \includegraphics[width = 0.9\textwidth]{FuelFairshare.png} \\ %565*238
  \caption*{\footnotesize \hspace{-1.6cm}$-2$ \hspace{0.5cm}$-1$ \hspace{0.5cm}$0$ \hspace{0.5cm}$1$ \hspace{0.5cm}$2$  }
      \end{columns}
    \end{figure}

\section{Conclusion}

\normalsize
\begin{frame}
\linespread{1.2}
\Wider[3.5em]{
\frametitle{Conclusion}
\begin{itemize}
\item The key factors associated with \textbf{climate knowledge}:
\begin{itemize}
\item Female: negative effect
\item Financial literacy: positive effect
\item Score on the Cognitive Reflection Test: positive effect 
\end{itemize}

\item The key factors associated with \textbf{WTP for climate change mitigation}:

\item Age: negative effect
\item Numeracy, financial literacy: negative effect
\item Attitudes and beliefs about redistribution of income and wealth
\item Climate change risk perception: positive effect (robustness check)
\end{itemize}

\end{itemize}}
\linespread{1}
\end{frame}

\note{ Interesting for climate knowledge: behavioral variabless not significant, not even opinions. education ordinal.pops out for the smaller sample Lasso \\
Interesting for climate knowledge: some evidencem that also climcare in 100 years, but not good for proportional odds thing\\
Discount rate mildly significant, SVO little bit significant for fuel\\
WTP: HOW MUCH FEMALE WANT TO PAY DOES NOT DEPEND ON CLIMPOL SO MUCH}

@@

\begin{frame}\label{ClimpolClimcare}
\Wider[0.9em]{
\frametitle{Climate change and policies risk perception}
\linespread{1.2}
\textbf{Climate care} \\
\vspace{0.3cm}
How serious a problem do you think climate change is at this moment?
\\

\vspace{0.3cm}
{\footnotesize

\begin{tabular}{lrrrrrrrrrrr} 
\hline
\rowcolor{PaleGreen}Response&$0$&$1$&$2$&$3$&$4$&$5$&$6$&$7$&$8$&$9$&$10$\\
 \hline
 \cellcolor{PaleGreen}Frequency&$77$&$68$&$112$&$171$&$199$&$430$&$858$&$1025$&$786$&$395$&$471$\\
\hline
\end{tabular}}

}
\Wider[1.8em]{
\vspace{0.6cm}
\textbf{Climate policy} \\
\vspace{0.3cm}
Which affects you and your way of life more, climate change (10) or policies (0) to reduce
greenhouse gas emissions?
\\

\vspace{0.3cm}
{\footnotesize

\begin{tabular}{lrrrrrrrrrrr} 
\hline
\rowcolor{PaleGreen}Response&$0$&$1$&$2$&$3$&$4$&$5$&$6$&$7$&$8$&$9$&$10$\\
 \hline
 \cellcolor{PaleGreen}Frequency&$130$&$127$&$256$&$374$&$415$&$1174$&$670$&$621$&$397$&$205$&$223$\\
\hline
\end{tabular}}
}

\vspace{0.6cm}
n=$4592$

\end{frame}
\note{The same for seriousness in 10 and 100 years and policy for children and grandchildren}

\begin{frame}\label{Data3}
\frametitle{Data} 
n=$4592$

 \begin{figure}
 \caption*{Occupation}
%\includegraphics[width = 1\textwidth]{Occupation.png}  
\end{figure}

\note{ hello}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Financial literacy} \\
\hspace{1cm} Number of correct answers out of three financial questions
\\
\vspace{0.3cm}
\begin{tabular}{lllllll} 
\hline
\multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{1.5} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{2.5} & \multicolumn{1}{c}{3}\\
\hline
\end{tabular}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Understands interest} \\
\hspace{1cm}1 = Understands compound interest
\\
\vspace{0.3cm}
\begin{tabular}{llll} 
\hline
\multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1} \\
\hline
\end{tabular}

\vspace{1.3cm}

\textbf{Understands inflation} \\
\hspace{1cm}1 = Understands inflation
\\
\vspace{0.3cm}
\begin{tabular}{cccc} 
\hline
\multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1} \\
\hline
\end{tabular}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Cognitive Reflection Test} \\
\begin{itemize}
 \item In survey called numeracy
 \item Frederick (2005)
 \item Number of correct answers - max is 3
\end{itemize}

\vspace{0.3cm}
\begin{tabular}{lllllll} 
\hline
\multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{1.5} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{2.5} & \multicolumn{1}{c}{3}\\
\hline
\end{tabular}

\vspace{1.3cm}

\note{The Cognitive reflection test measures the ability or disposition to reflect on a question and resist reporting the first response that comes to mind}

\begin{frame}\label{Questions}

>>>>

\title{Drought and its Effects on Economy}  
\author{Monika Novackova} 
\date{December 2017} 
\institute{University of Sussex}

@@

\maketitle

\begin{frame}\label{My Interest}

\large{My research interest:} \\
\vspace{0.7cm}

\large{\hspace{1cm}\textbf{Effects of droughts on economy in Kenya}}
\end{frame}

@@

\section{Definition of Drought}

@@

\section{Drought Indices}

@@

 \begin{itemize}
 \item \textbf{Standardized Precipitation Evapotranspiration Index} (SPEI, \citealp{SPEI})
\begin{itemize}
\item Extention of SPI
\item Accounts for potential evapotranspiration (hence captures impacts of increased temperature on water demand)
\end{itemize}
\item Number of other drought indicators and indices exist
\begin{itemize}
\item E.g.: Percent of Normal Precipitation, Drought Area Index (DAI), Soil Moisture Anomaly (SMA), Standardized Water-level Index (SWI), Normalized Difference Vegetation Index (NDVI)
\end{itemize}
\item  \textit{For a detailed overview of drought indices see e.g.:} \cite{monacelli2005}, \cite{svoboda2016} or \cite{zargar2011}

\item \cite{keyantash2002}: The quantification of drought: an evaluation of drought indices
\end{itemize}

@@

\section{Effects of Droughts}

@@

\frametitle{Effects of droughts on economy}\label{Effects} 
\begin{block}{\textbf{Computable General Equilibrium (CGE)} Models }
\begin{itemize}

\item Consists of equations describing model variables and a database (input-output tables, SAM matrix)
\item Assuming optimizing behaviour (cost minimizing producers, optimal households demands)

\item \underline{\textbf{\cite{OxfamIDS}}}
\begin{itemize}
\item Exploring range of scenarios for food price increase in 2030
\begin{itemize}
\item 1. Baseline 2. Climate change 3. Climate change with adaptation 4. Adaptation only in sub-Saharan Africa
\end{itemize}
\item Global coverage, set of individual country models, linked through international trade
\item Climate change (incl. drought) modelled as changes in factor productivity (usually negative)

\end{itemize}
\end{block}
\end{frame}

\frametitle{Effects of droughts on economy}\label{Effects} 
\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
\begin{itemize}

\item[~] \underline{\textbf{\cite{OxfamIDS}}}
\begin{itemize}
\item Solves the within-country models and between-country trade
relationships simultaneously
\item 19 region, 12 sector/commodity group, 8 commodity groups
\end{itemize}
\begin{table}
\begin{footnotesize}
\begin{tabular}{lll} 
\hline
\rowcolor{PaleGreen}  & \textbf{Determined by model}  \\
\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{(outputs)}  \\
 \hline
Agricultural productivity growth	& Production volumes of \\
&  \hspace{0.5cm}food commodities \\
Commodity shares in hh. expenditure &Production vol. per capita \\
Shares of import in demand (commodities) & World market  \\
GDP growth rates&\hspace{0.5cm} food prices (change)\\
Population growth& Domestic food price (change)\\
& Volumes of global trade\\
Shares of food in hh. expenditure  & Aggregate index of
\\
 & \hspace{0.5cm}consumption p.capita
\\
 \hline
\end{tabular}
\end{footnotesize}
\end{table}

\end{itemize}
\end{block}
\end{frame}

\note{\\
\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
\vspace{0.3cm}a\\
\vspace{0.3cm}
\vspace{0.3cm}
}

\frametitle{Effects of droughts on economy}\label{Effects} 
\begin{block}{\textbf{Computable General Equilibrium (CGE)}}

\begin{itemize}
\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item Ethiopia, Social Accounting Matrix (SAM)

\item Model drought as a $20\%$ reduction in crop productivity and $20$ reduction of livestock capital
\item Effects of shocks in production on cereal prices and food consumption
\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
\end{itemize}
\end{itemize}
\end{block}
\end{frame}

\note{
\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP. Also models price transition between international and domestic prices..
\\
\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

\frametitle{Effects of droughts on economy}\label{Effects} 
\begin{block}{\textbf{Computable General Equilibrium (CGE)}}

\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item 5 agro-ecological zones,46 production activities (incl. 35 zone specific agricultural production sectors), 22 commodity groups,  15 primary factors of production
\end{itemize}

\begin{table}
\begin{footnotesize}
\begin{tabular}{lll}

\hline
\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{Determined by model (outputs)}  \\
 \hline
Capital stock& Domestic price of each commodity \\
Land (by region) & Land allocated across crops \\
Supply of labor per skill type& Real wages\\
Foreign capital inflow	& Real exchange rate\\
Trade balance	& \\
 \hline
\end{tabular}
\end{footnotesize}
\end{table}

\begin{footnotesize}

\item The simulation use a 'balanced' macro closure in which aggregate investment, government demand, and consumption are fixed shares of total absorption
\item Intermediate inputs into production are determined as fixed shares
of the quantity of output

\end{footnotesize}
\end{itemize}
\end{block}
\end{frame}
\note{\\

\vspace{0.3cm}
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

\frametitle{Effects of droughts on economy}\label{Effects}

\item \textbf{\cite{Pedram2016}}

\begin{itemize}
\item National production losses per disaster (droughts, floods and extreme temperatures)
\item Worldwide, superposed epoch analysis
\item On average $10.1\%$ reduction of cereal production can be linked to drought
\end{itemize}
\item \textbf{\cite{Mehrabi2017}}
\begin{itemize}
\item Cumulative production losses linked to extreme heat and drought events
\item Per years (53), countries (131), commodities (6)
\end{itemize}

\end{frame}
\note{Lesk: 9.1 percent reduction is attributable to extreme heat.
They did not find any significant effect of extreme cold and 
oods on production.

\vspace{0.3cm} mehrabi: estimate counter factual production in disaster years
\vspace{0.3cm}mehrabi:units are a bit like a panel - years x commodity x country x disaster (disaster years and not disaster years
\\mehrabi: difference in production of disaster years and non-disaster year average. Cumulative impact: sum per countries and disasters. Loss relatively to (divided by) production + loss of production
\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

\section{Effects of Climate}

\frametitle{Effects of climate and weather on economy}\label{EffectsCLimate}

\item Effects on food prices

\item \textbf{\cite{Brown2015}}
\begin{itemize}
\item  Climate measured by NDVI vegetation index 
\item  Effects of NDVI and world food prices on local food prices
\item $20\%$ prices affected by weather, $9\%$ by international prices
\end{itemize}

\item Effects on production (production function approach)
\begin{itemize}
\item \textbf{\cite{Ochieng2016}}
\begin{itemize}
\item Effect of climate variability and change on crop revenue
\item Augmented production function, Kenya
\item  Household FE
\end{itemize}
\item \textbf{\cite{Deschenes2007Ric}:} US, positive effect of global warming
\end{itemize}
\item Effects on land prices or yields/acre (Ricardian analysis)
\begin{itemize}
\item  \textbf{\citet{kabubo2007}:} Kenya, warming harmful

\item  \cite{SeoMendelsohn} and \cite{KMendelsohn2008}: mild and wet warming - positive, dry more severe warming - negative

\end{itemize}
\end{frame}
\note{brown, Khsingar - KALMAN FILTER

\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	
\\
\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}
% tady mozna pridat nejake ricardian studie a rozdelit to na dva slidy

\section{My Idea}

\frametitle{My idea - panel estimation}\label{MySuggestion}

My interest: \textbf{Effects of drought on economy in Kenya}
\begin{itemize}

\item \textbf{Outcome variable}

\begin{itemize}
\item Volumes of production (crop specific, total)
\item Alternatively food prices, yields/per acre
\end{itemize}

\item \textbf{Units of analysis}
\begin{itemize}
\item Counties in Kenya $\times$ year
\
\end{itemize}

\item \textbf{Explanatory variable of interest}

\begin{itemize}
\item Dummy variable (0/1) drought occurred in a particular county and year or not
\item Several varieties - various specifications of drought:

@@

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

@@

\frametitle{My idea - panel estimation}\label{MySuggestion}

My interest: \textbf{Effects of drought on economy in Kenya}
\begin{itemize}

\item \textbf{Explanatory variable of interest}

\begin{itemize}
\item Dummy variable (0/1) drought occurred in a particular county and year or not
\item Several varieties - various specifications of drought:

\begin{itemize}
\item Drought index monthly - NDVI, SPI, SPEI 
\item Drought index below a specific value ($-1$, $-2$ or $-3$)
 \begin{description}[abc]  % for indentation of length of abc
 \item[$\bullet$] At least for one month during a specific year
 \item[$\bullet$] At least for one month during growing season in a specific year
 \item[$\bullet$] At least for one month during long rains in a specific year
  \item[$\bullet$] At least for two (three) consecutive months during a growing season in a specific year
 \item[$\bullet$] At least for one month during two (three) consecutive growing seasons
 \end{description}

\end{itemize}
\end{itemize}

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

\frametitle{My idea - panel estimation}\label{MySuggestion}

My interest: \textbf{Effects of drought on economy in Kenya}
\begin{itemize}

\item \textbf{Explanatory variable of interest}

\begin{itemize}
\item Dummy variable (0/1) drought occurred in a particular county and year or not

\item \textbf{Units of analysis:} Counties in Kenya $\times$ year

\item \textbf{Control variables ?}

\begin{itemize}
\item GDP, population, land area, soil characteristics, climate (2 seasons)?
\end{itemize}

\item \textbf{Estimation methods}

\begin{itemize}
\item Fixed effects, SURE, Kalman filter, Box-Jenkins
\end{itemize}

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

@@

\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc o}

@@

@@

\frame[allowframebreaks]{ 
\small
\frametitle{References}
\bibliographystyle{apa}
\bibliography{referencesFS}
}
